Efficient Handling of String-Number Conversion

### 2020-04-02 draft.tex
small fixes related to rev C and few other minor things
change running example

For ease of presentation, we  use the following toy example $$\Phi = \{xy=yx, n = \sti{x}, n >3 ,|y|>|x|, y \in (12)^+\}$$

to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that $x$ concatenated with $y$ is the same as $y$ with $x$, $n$ is the numeric value of the string $x$, $n > 3$, $y$ is longer than $x$, and $y$ is a string in the form of a finite repetitions of $``12"$.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12"$ and $y=``1212"$.

@@

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
Let $Z$ be the set of integer variables in $\insc$.
The function maps an interpretations $I_e$ over $Z\cup V_\dr\cup\#V_\dr$
to an interpretation over $Z \cup X$, following the domain restriction $\dr$. Informally, it ``decodes" an interpretation of integer variables $Z\cup V_\dr\cup\#V_\dr$ to an interpretation of variables in the string constraint $\insc$. Formally,
we define $\decode  \dr (I_e) \defeq \{I \mid \forall z\in Z: I(z) = I_e(z) \wedge\forall x\in X: \{I(x)\} = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that (1) $I$ and $I_e$ are consistent over variables in $Z$ and (2) $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$ is a word encoding that $\dr(x)$-encodes $I(x)$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a interpretation $I$ of the string constraint $\insc$,  
we let $\encode \dr (I) =\{I_e\mid \decode \dr (I_e) = \{I\} \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{example}
We consider the domain restriction $\dr$ such that $\dr(x)	 =(A_x,\true)$ from Figure~\ref{fig:sfa}~(a) and $\dr(y) = (A_y,\true)$ from Figure~\ref{fig:sfa}~(b). 
%We assume that the set of integer variables is $Z=\{v_z\}$ and $V_{\dr(x)} =\{v_0, v_1,v_2,v_3\}$. 
Let the set of integer variables be $Z=\{v_z\}$ and let $V_{\dr(x)} =\{v_0, v_1,v_2,v_3\}$. 
For the interpretation $I_e =(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3) \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2)$, we have $\decode \dr (I_e) = \{(z,x,y) \rightarrow(3, ``ababab",``cc") \}$.\\ 
Conversely, for $I=(z,x,y) \rightarrow(3, ``ababab",``cc")$ and $Y=(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3)$, we have $\encode \dr (I) =\left\{ 
\begin{array}{l}
Y \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2), \\
Y \rightarrow (3,\semof{a}, \semof{b},\semof{\epsilon}, \semof{c},3,3,2,2), \\
Y\rightarrow (3,\semof{a}, \semof{b},\semof{c},\semof{c},3,3,1,1)
\end{array}
 \right\}$
\end{example}

>>>>

For ease of presentation, we  use the following toy example $$\Phi = \{`0'x = x`0', \sti{x} = \sti{y} ,|y|>|x|, 1000<|y| \}$$

to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that $`0'$ concatenated with $x$ is the same as $x$ with $`0'$, $n$ is the numeric value of the string $x$, $n > 3$, $y$ is longer than $x$, and $y$ is a string in the form of a finite repetitions of $``12"$.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12"$ and $y=``1212"$.

@@

The correctness of the entire construction of $\underf \dr {\insc}$ is expressed by  Theorem~\ref{theorem:correct}. 
It uses the decoding function $\decode \dr$ parameterized by the domain restriction $\dr$. 
Let $Z$ be the set of integer variables in $\insc$.
The function maps an interpretations $I_e$ over $Z\cup V_\dr\cup\#V_\dr$
to an interpretation over $Z \cup X$, following the domain restriction $\dr$. Informally, it ``decodes" an interpretation of integer variables $Z\cup V_\dr\cup\#V_\dr$ to an interpretation of variables in the string constraint $\insc$. Formally,
we define $\decode  \dr (I_e) \defeq \{I \mid \forall z\in Z: I(z) = I_e(z) \wedge\forall x\in X: \{I(x)\} = \decode {\dr(x)} ((I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}})\}$.
The condition says that (1) $I$ and $I_e$ are consistent over variables in $Z$ and (2) $(I_e)_{V_{\dr(x)}\cup \#V_{\dr(x)}}$ is a word encoding that $\dr(x)$-encodes $I(x)$.
%We define the output of $\decode  \dr (I_e\cup I_z)$ as a set so that \emph{invalid word encodings}, i.e. those do not corresponds to possible word encodings from PFAs, are mapped to empty sets and valid ones are mapped to singleton sets.
We also define the $\dr$-encoding function as the counterpart of $\dr$-decoding,  
%
namely, for a interpretation $I$ of the string constraint $\insc$,  
we let $\encode \dr (I) =\{I_e\mid \decode \dr (I_e) = \{I\} \}$.
We lift $\decode \dr$ and $\encode \dr$ to sets of interpretations in the standard manner.
\begin{example}
We consider the domain restriction $\dr$ such that $\dr(x)	 =(A_x,\true)$ from Figure~\ref{fig:sfa}~(a) and $\dr(y) = (A_y,\true)$ from Figure~\ref{fig:sfa}~(b). 
%We assume that the set of integer variables is $Z=\{v_z\}$ and $V_{\dr(x)} =\{v_0, v_1,v_2,v_3\}$. 
For the purpose of this example, let the set of integer variables be $Z=\{v_z\}$. 
We have $V_{\dr(x)} =\{v_0, v_1,v_2,v_3\}$. 
For the interpretation $I_e =(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3) \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2)$, we have $\decode \dr (I_e) = \{(z,x,y) \rightarrow(3, ``ababab",``cc") \}$.\\ 
Conversely, for $I=(z,x,y) \rightarrow(3, ``ababab",``cc")$ and $Y=(v_z,v_0,v_1,v_2,v_3,\#v_0,\#v_1,\#v_2,\#v_3)$, we have $\encode \dr (I) =\left\{ 
\begin{array}{l}
Y \rightarrow (3,\semof{a}, \semof{b}, \semof{c},\semof{\epsilon},3,3,2,2), \\
Y \rightarrow (3,\semof{a}, \semof{b},\semof{\epsilon}, \semof{c},3,3,2,2), \\
Y\rightarrow (3,\semof{a}, \semof{b},\semof{c},\semof{c},3,3,1,1)
\end{array}
 \right\}$
\end{example}
